MCC Ex3 Variable Priors

### 2019-03-29 papers/full_paper/SI_appendix.Rmd
first draft results and methods

@@

ex_3_pred <- bind_rows(
  ex_3_model_pred,
  ex3_pragm_model_pred_noise,
  adult_ex3_data_summary
)

@@

```{r plot ex3 model predictions and data adults}
ggplot(data = ex_3_pred, aes(x = alignment, y = mean, col = model)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), size = .4,  position = position_dodge(width = .5), pch = 5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Choosing More Informative")+
  facet_grid (~ common_ground_manipulation + speaker) +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_color_manual(name="Model",
                     breaks=c("data","pragmatic","pragmatic_noise","flat_prior_noise","prior_only_noise"),
                     values= c("black","#859900","#00afaf","#6c71c4","#b58900"))+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))
```

@@

#### Loglike Parameter free model
```{r log likes ex3 adults}
# calculating model log-likelihoods
log_like_pragmatic_ex3 <- adult_ex3_data %>%
  left_join(ex3_pragm_model_predictions)%>%
  mutate(like = ifelse(correct_inf == 1, value, 1 - value),
         log_like = log(like))%>%
  group_by(model)%>%
  summarise(logP = sum(log_like))
```

#### Loglike models with noise

##### Pragmatic model 
```{r model rsa noise ex3, include = FALSE}
loglike.rsa.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

@@

#### Bayes factors

@@

@@

```{r prior pre-tests data adults}
# load data from prior strength manipulation experiments to set priors
adult_ex4_prior_data <-read_csv(file="../../stats/data/adult_ex4_prior.csv") 
```

@@

strongPrior <- expand.grid(
  common_ground_manipulation = c("preference","novelty"),
  speaker = c("same_speaker","different_speaker"),
  prior_manipulation = c("strong", "medium", "weak")
)

strong_prior_summary <- adult_ex4_prior_data %>%
  group_by(common_ground_manipulation, prior_manipulation, speaker) %>%
  summarize(prior = mean(correct))

model_params_ex4 <- left_join( 
  strongPrior, 
  strong_prior_summary# %>% filter(prior_manipulation == "strong")
) %>%
  rename(congruent = prior) %>%
  mutate(incongruent = 1 - congruent) %>%
  gather(alignment, prior, congruent, incongruent) %>%
  drop_na() %>%
  mutate(speak_opt = adult_model_param$so_map)
```

@@

```{r plot ex4 model predictions and data adults}
ggplot(data = ex_4_pred, aes(x = alignment, y = mean, col = model)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), size = .4,  position = position_dodge(width = .5), pch = 5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Choosing More Informative")+
  facet_grid ( prior_manipulation ~ common_ground_manipulation + speaker) +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_color_manual(name="Model",
                     breaks=c("data","pragmatic_noise","flat_prior_noise","prior_only_noise"),
                     values= c("black","#859900","#6c71c4","#b58900"))+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))
```

@@

```{r plot for loglikelihood ex4 adults}
ggplot(ex4_model_comparison, aes(x = model, y = logP, col = model))+
  geom_segment( aes(x=model, xend=model, y=logP, yend=0), col = "black", size = 1)+
  geom_point(size = 7)+
  theme_few()+ 
  xlab("Model")+
  ylab("Log-likelihood")+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  guides(col = F)+
  scale_color_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#b58900","#859900","#6c71c4"))
```

#### Bayes factors

@@

```{r model setup ex3 children}
ex3_model_data <- child_ex3_data%>%
  mutate(so_int = child_model_param$so_int,
         so_slope = child_model_param$so_slope,
         pref_int_different_speaker = child_model_param$pref_int_different_speaker,
         pref_slope_different_speaker = child_model_param$pref_slope_different_speaker,
         pref_int_same_speaker = child_model_param$pref_int_same_speaker,
         pref_slope_same_speaker = child_model_param$pref_slope_same_speaker,
         so = so_int+so_slope*(age_num-min(age_num)),
         diff_prior = plogis(pref_int_different_speaker+pref_slope_different_speaker*(age_num-min(age_num))),
         same_prior = plogis(pref_int_same_speaker+pref_slope_same_speaker*(age_num-min(age_num))))%>%
  group_by(id,trial)%>%
        mutate(priorProbs = ifelse(speaker == "same_speaker" & alignment == "congruent", list(c(1-same_prior,1-same_prior,same_prior)),
                                   ifelse(speaker == "different_speaker" & alignment == "congruent", list(c(1-diff_prior,1-diff_prior,diff_prior)),
                                          ifelse(speaker == "same_speaker" & alignment == "incongruent", list(c(same_prior,same_prior,1-same_prior)),
                                                 list(c(diff_prior,diff_prior,1-diff_prior))

                                   ))))%>%
  ungroup()%>%
  select(speaker,age_num,minage,alignment,so,priorProbs,correct_inf)

@@

```{r plot model predictions ex3 children}
child_model_predictions <- bind_rows(
  child_ex3_pragm_model_predictions_noise,
  #child_ex3_pragm_model_predictions_developmental_noise,
  child_ex3_pragm_model_predictions,
  child_ex3_prior_only_model_predictions_noise,
  child_ex3_flat_prior_model_predictions_noise
  #child_ex3_flat_prior_model_predictions,
  #child_ex3_prior_only_model_predictions
  )

@@

ggplot(child_model_predictions)+
  #geom_jitter(data = plot_child_ex3_data, aes(x = age_num, y = mean), width = .025,height = .025, alpha = .2)+
  geom_ribbon(aes(x = age, ymin = ci_lower, ymax = ci_upper, fill = model), alpha = .4) +
  geom_line(aes(x = age, y = mean, col = model), size = 1, alpha = 1, linetype = 1) + 
  geom_smooth(data = child_ex3_data, aes(x = age_num, y = correct_inf), col = "black", size = 0.6, method = "loess", span = 2,  se = T,  alpha = .5)+
  #geom_hline(yintercept = 0.5, lty=2)+
  labs(y="proportion more informative")+
  ylim(-0.05,1.05)+
  facet_grid(speaker~alignment)+
  theme_few()+
  scale_color_manual(name="Model",
                     breaks=c("pragmatic","pragmatic_noise","flat_prior_noise","prior_only_noise"),
                     values= c("#b58900","#00afaf","#6c71c4","#859900"))+
  scale_fill_manual(name="Model",
                     breaks=c("pragmatic","pragmatic_noise","flat_prior_noise","prior_only_noise"),
                     values= c("#b58900","#00afaf","#6c71c4","#859900"))

@@

```{r correlation plot ex3 children}
child_cor_data <- child_ex3_data %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by(speaker, alignment, age_bin) %>%
  summarize(k = sum(correct_inf), n = n())%>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         Data_ci_lower  = qbeta(.025, a, b),
         Data_ci_upper = qbeta(.975, a, b),
         Data = (a-1)/(a+b-2))%>%
  select(-a,-b,-n,-k)

child_cor_model <- child_model_predictions %>%
  mutate(age = ifelse(age<4,"3","4"))%>%
  group_by(model,speaker,alignment, age)%>%
  summarise(Model = mean(mean))

child_cor_plot <- child_cor_model %>%
  ungroup()%>%
  mutate(Data = rep(child_cor_data$Data,4),
         Data_ci_lower = rep(child_cor_data$Data_ci_lower,4),
         Data_ci_upper  =rep( child_cor_data$Data_ci_upper,4))

ggplot(data = child_cor_plot,aes(x = Model, y = Data, col = age)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_point(size = 3)+
  geom_errorbar(aes(ymin = Data_ci_lower, ymax = Data_ci_upper),width = 0,size = .5)+
  coord_fixed()+
  facet_grid(~model)+
  stat_cor(method = "pearson", label.x = 0.01, label.y = 0.99, aes(x = Model, y = Data), inherit.aes = F)+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  theme_few() + 
  scale_colour_ptol()
```

@@

child_ex3_model_comparison <- bind_rows(
  child_ex3_pragm_model_likelihood,
  child_ex3_pragm_model_noise_likelihood,
  child_ex3_pragm_model_developmental_noise_likelihood,
  child_ex3_flat_prior_model_noise_likelihood,
  child_ex3_flat_prior_model_likelihood,
  child_ex3_flat_prior_model_developmental_noise_likelihood,
  child_ex3_prior_only_model_noise_likelihood,
  child_ex3_prior_only_model_likelihood,
  child_ex3_prior_only_model_developmental_noise_likelihood
)%>%
  ungroup()%>%
  mutate(model = reorder(model,logP))
```

@@

### Bayes factors

>>>>

We noted that the model predictions were relatively

@@

ex_3_pred <- bind_rows(
  ex_3_model_pred,
  adult_ex3_data_summary
)

@@

```{r plot ex3 model predictions and data adults}
ggplot(data = ex_3_pred, aes(x = alignment, y = mean, col = model)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), size = .4,  position = position_dodge(width = .5), pch = 5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion more informative")+
  facet_grid (~ common_ground_manipulation + speaker) +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_color_manual(name="Model",
                     breaks=c("data","pragmatic","pragmatic_noise","flat_prior_noise","prior_only_noise"),
                     values= c("black","#859900","#00afaf","#6c71c4","#b58900"))+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))
```

@@

##### Pragmatic model

###### Parameter free

```{r log likes ex3 adults}
# calculating model log-likelihoods
# log_like_pragmatic_ex3 <- adult_ex3_data %>%
#   left_join(ex3_pragm_model_predictions)%>%
#   mutate(like = ifelse(correct_inf == 1, value, 1 - value),
#          log_like = log(like))%>%
#   group_by(model)%>%
#   summarise(logP = sum(log_like))%>%
#   saveRDS("../../stats/saves/ex3_pragm_model_loglike.rds")

log_like_pragmatic_ex3 <- readRDS("../../stats/saves/ex3_pragm_model_loglike.rds")
```

###### Noise parameter

```{r model rsa noise ex3, include = FALSE}
loglike.rsa.ex3.noise <-'
var allParams = dataFromR.allParams
var allData = dataFromR.allData

@@

###### Noise parameter

@@

#### Log Bayes factors

@@

### Prior manipulation experiments

```{r ex4 prior pre-tests data adults}
# load data from prior strength manipulation experiments to set priors
adult_ex4_prior_data <-read_csv(file="../../stats/data/adult_ex4_prior.csv") 
```

```{r ex4 prior pre-test plots}
p_adult_ex4_priors <- adult_ex4_prior_data%>%
  mutate(prior_manipulation = relevel(as.factor(prior_manipulation), ref = "strong"))%>%
  group_by(common_ground_manipulation,prior_manipulation,speaker, id) %>%
  summarise(correct = mean(correct))

p_adult_ex4_priors_ci <- p_adult_ex4_priors %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p_adult_ex4_priors, aes(x = speaker, y = correct, col = speaker), alpha = 0.3, width = .3,height = .02)+
  geom_pointrange(data = p_adult_ex4_priors_ci, aes(x = speaker, y = mean, col = speaker,ymin = ci_lower, ymax = ci_upper),size = .8, pch = 5, fatten = 2.5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="")+
  facet_grid(common_ground_manipulation ~ prior_manipulation)+
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F,size =F)+
  scale_color_solarized()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

@@

@@

ex4Prior <- expand.grid(
  common_ground_manipulation = c("preference","novelty"),
  speaker = c("same_speaker","different_speaker"),
  prior_manipulation = c("strong", "medium", "weak")
)

ex4_prior_summary <- adult_ex4_prior_data %>%
  group_by(common_ground_manipulation, prior_manipulation, speaker) %>%
  summarize(prior = mean(correct))

model_params_ex4 <- left_join( 
  ex4Prior, 
  ex4_prior_summary
) %>%
  rename(congruent = prior) %>%
  mutate(incongruent = 1 - congruent) %>%
  gather(alignment, prior, congruent, incongruent) %>%
  drop_na() %>%
  mutate(speak_opt = adult_model_param$so_map)
```

@@

```{r plot ex4 model predictions and data adults}
ggplot(data = ex_4_pred, aes(x = alignment, y = mean, col = model)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), size = .4,  position = position_dodge(width = .5), pch = 5)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion more informative")+
  facet_grid ( prior_manipulation ~ common_ground_manipulation + speaker) +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_color_manual(name="Model",
                     breaks=c("data","pragmatic_noise","flat_prior_noise","prior_only_noise"),
                     values= c("black","#b58900","#6c71c4","#859900"))+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))

@@

```{r plot for loglikelihood ex4 adults}
ggplot(ex4_model_comparison, aes(x = model, y = logP, col = model))+
  geom_segment( aes(x=model, xend=model, y=logP, yend=0), col = "black", size = 1)+
  geom_point(size = 7)+
  theme_few()+ 
  xlab("Model")+
  ylab("Log-likelihood")+
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  guides(col = F)+
  scale_color_manual(breaks=c("pragmatic","prior_only","flat_prior"),
                    values= c("#859900","#b58900","#6c71c4"))
```

#### Log Bayes factors

@@

```{r model setup ex3 children}
ex3_model_data <- child_ex3_data%>%
  mutate(so_int = child_model_param$so_int,
         so_slope = child_model_param$so_slope,
         pref_int_different_speaker = child_model_param$pref_int_different_speaker,
         pref_slope_different_speaker = child_model_param$pref_slope_different_speaker,
         pref_int_same_speaker = child_model_param$pref_int_same_speaker,
         pref_slope_same_speaker = child_model_param$pref_slope_same_speaker,
         so = so_int+so_slope*(age_num-min(age_num)),
         diff_prior = plogis(pref_int_different_speaker+pref_slope_different_speaker*(age_num-min(age_num))),
         same_prior = plogis(pref_int_same_speaker+pref_slope_same_speaker*(age_num-min(age_num))))%>%
  group_by(id,trial)%>%
        mutate(priorProbs = ifelse(speaker == "same_speaker" & alignment == "congruent", 
                                   list(c(1-same_prior,1-same_prior,same_prior)),
                                   ifelse(speaker == "different_speaker" & alignment == "congruent",
                                          list(c(1-diff_prior,1-diff_prior,diff_prior)),
                                          ifelse(speaker == "same_speaker" & alignment == "incongruent",
                                                 list(c(same_prior,same_prior,1-same_prior)),
                                                 list(c(diff_prior,diff_prior,1-diff_prior))
                                   ))))%>%
  ungroup()%>%
  select(speaker,age_num,minage,alignment,so,priorProbs,correct_inf)

@@

```{r plot model predictions ex3 children}
child_model_predictions <- bind_rows(
  child_ex3_pragm_model_predictions_noise,
  child_ex3_pragm_model_predictions_developmental_noise, 
  child_ex3_pragm_model_predictions
  #child_ex3_prior_only_model_predictions_noise,
  #child_ex3_flat_prior_model_predictions_noise
  #child_ex3_flat_prior_model_predictions,
  #child_ex3_prior_only_model_predictions
  )

@@

ggplot(child_model_predictions)+
  geom_jitter(data = plot_child_ex3_data, aes(x = age_num, y = mean), width = .025,height = .025, alpha = .2)+
  geom_ribbon(aes(x = age, ymin = ci_lower, ymax = ci_upper, fill = model), alpha = .4) +
  geom_line(aes(x = age, y = mean, col = model), size = 1, alpha = 1, linetype = 1) + 
  #geom_smooth(data = child_ex3_data, aes(x = age_num, y = correct_inf), col = "black", size = 0.6, method = "loess", span = 2,  se = T,  alpha = .5)+
  #geom_hline(yintercept = 0.5, lty=2)+
  labs(y="Proportion more informative")+
  ylim(-0.05,1.05)+
  facet_grid(~speaker+alignment)+
  theme_few()+
  theme(legend.position="bottom")+
  scale_color_manual(name="Model",
                     breaks=c("pragmatic","pragmatic_noise","pragmatic_developmental_noise"),
                     values= c("#00afaf","#af6300","#6c71c4"))+
  scale_fill_manual(name="Model",
                     breaks=c("pragmatic","pragmatic_noise","pragmatic_developmental_noise"),
                     values= c("#00afaf","#af6300","#6c71c4"))

@@

```{r correlation plot ex3 children}
child_cor_data <- child_ex3_data %>%
  group_by(age_bin,speaker, alignment) %>%
  summarize(k = sum(correct_inf), n = n())%>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         data_ci_lower  = qbeta(.025, a, b),
         data_ci_upper = qbeta(.975, a, b),
         data_mode = (a-1)/(a+b-2))%>%
  select(-a,-b,-n,-k)

child_cor_model <- bind_rows(
  child_ex3_pragm_model %>%
  separate(`0`, into = c("age", "speaker", "alignment","correct"), sep="/")%>%
  mutate(value = `1`,
         model = "pragmatic_parameter_free",
         age = ifelse(age<4,"3","4"))%>%
    group_by(age,model,speaker,alignment)%>%
    summarise(model_mode = estimate_mode(value),
              model_ci_lower = hdi_lower(value), 
              model_ci_upper =hdi_upper(value)),
  readRDS("../../stats/saves/child_ex3_pragm_model_noise.rds") %>%
    filter(!(Parameter %in% c("noise"))) %>%
    separate(Parameter, into = c("age", "speaker", "alignment"), sep="/")%>%
    mutate(model = "pragmatic_noise",
           age = ifelse(age<4,"3","4"))%>%
    group_by(age,model,speaker,alignment) %>%
    summarise(model_mode = estimate_mode(value), 
              model_ci_lower = hdi_lower(value), 
              model_ci_upper =hdi_upper(value)),
  readRDS("../../stats/saves/child_ex3_pragm_model_developmental_noise.rds") %>%
    filter(!(Parameter %in% c("noise","noise_int","noise_slope"))) %>%
    separate(Parameter, into = c("age", "speaker", "alignment"), sep="/")%>%
    mutate(model = "pragmatic_developmental_noise",
           age = ifelse(age<4,"3","4"))%>%
    group_by(age,model,speaker,alignment) %>%
    summarise(model_mode = estimate_mode(value), 
              model_ci_lower = hdi_lower(value), 
              model_ci_upper =hdi_upper(value))
)

child_cor_plot <- child_cor_model %>%
  ungroup()%>%
  mutate(data_mode = rep(child_cor_data$data_mode,3),
         data_ci_lower = rep(child_cor_data$data_ci_lower,3),
         data_ci_upper  =rep( child_cor_data$data_ci_upper,3))

ggplot(data = child_cor_plot,aes(x = model_mode, y = data_mode, col = age)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_point(size = 3)+
  geom_errorbar(aes(ymin = data_ci_lower, ymax = data_ci_upper),width = 0,size = .7)+
  geom_errorbarh(aes(xmin = model_ci_lower, xmax = model_ci_upper), height = 0,size = .7)+
  coord_fixed()+
  facet_grid(~model)+
  stat_cor(method = "pearson", label.x = 0.01, label.y = 0.99, aes(x = model_mode, y = data_mode), inherit.aes = F)+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  theme_few() + 
  scale_colour_ptol()
```

@@

# child_ex3_model_comparison <- bind_rows(
#   child_ex3_pragm_model_likelihood,
#   child_ex3_pragm_model_noise_likelihood,
#   child_ex3_pragm_model_developmental_noise_likelihood,
#   child_ex3_flat_prior_model_noise_likelihood,
#   child_ex3_flat_prior_model_likelihood,
#   child_ex3_flat_prior_model_developmental_noise_likelihood,
#   child_ex3_prior_only_model_noise_likelihood,
#   child_ex3_prior_only_model_likelihood,
#   child_ex3_prior_only_model_developmental_noise_likelihood
# )%>%
#   ungroup()%>%
#   mutate(model = reorder(model,logP))%>%
#   saveRDS("../../stats/saves/child_ex3_model_comparison.rds")

child_ex3_model_comparison <- readRDS("../../stats/saves/child_ex3_model_comparison.rds")

@@

### Log Bayes factors

```{r}
child_ex3_pragmatic_models_bf <- child_ex3_model_comparison%>%
  filter(model == "pragmatic")%>%
  spread(parameter, logP) %>%
  mutate("dev. noise > noise" = `developmental noise` - noise,
         "noise > parameter free" = noise - `parameter free`,
         "dev. noise > parameter free" = `developmental noise` - `parameter free`) %>%
  select(-`parameter free`,-`developmental noise`, -noise)

child_ex3_pragmatic_models_bf%>%
  kable(digits = 2)
```

### 2019-03-29 stats/adults/ex3_2_variable_priors_comparison.Rmd
first draft results and methods

dat <- bind_rows(read_csv(file="ex3.2.novel.strong.data.csv"),
               read_csv(file="ex3.2.novel.medium.data.csv"),
               read_csv(file="ex3.2.novel.weak.data.csv"),
               read_csv(file="ex3.2.pref.strong.data.csv"),
               read_csv(file="ex3.2.pref.medium.data.csv")
               ) %>%
  mutate(trial_type = ifelse(trial == "train1" | trial =="train2", "train", "test"))
```

@@

m.data.strong <- data %>%
  filter(Prior == "Strong") %>%
  mutate(item = ifelse( change == "false", agent, altAgent))

@@

m.data.medium <- data %>%
  filter(Prior == "Medium") %>%
  mutate(item = ifelse( change == "false", agent, altAgent))

@@

m.data.weak <- data %>%
  filter(Prior == "Weak") %>%
  mutate(item = ifelse( change == "false", agent, altAgent))

>>>>

dat <- bind_rows(read_csv(file="../data/ex3.2.novel.strong.data.csv"),
               read_csv(file="../data/ex3.2.novel.medium.data.csv"),
               read_csv(file="../data/ex3.2.novel.weak.data.csv"),
               read_csv(file="../data/ex3.2.pref.strong.data.csv"),
               read_csv(file="../data/ex3.2.pref.medium.data.csv")
               ) %>%
  mutate(trial_type = ifelse(trial == "train1" | trial =="train2", "train", "test"))
```

@@

m.data.strong <- data %>%
  filter(Prior == "Strong") %>%
  mutate(item = ifelse( change == "same", agent, altAgent))

@@

m.data.medium <- data %>%
  filter(Prior == "Medium") %>%
  mutate(item = ifelse( change == "same", agent, altAgent))

@@

m.data.weak <- data %>%
  filter(Prior == "Weak") %>%
  mutate(item = ifelse( change == "same", agent, altAgent))
